Project-I by Group Sydney

### 2015-10-31 report/paper.tex
Finish report

Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable. Moreover, the results obtained were less good compared to the ridge regression method, and so we will not report those. The method using the least squares with gradient descent is appropriated as well as ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix. It is also faster and a little bit more efficient than least squares using gradient descent and so, only results obtained with this latter will be reported.

@@

The sixth model consists of using dummy encoding for the categorical variables. Unfortunately, we didn't get major improvements. We think that the reason why dummy encoding doesn't help is because we don't have a high bias and so, having more features doesn't help.

@@

For the seventh model, the degrees of the polynomial for the clusters are $3,6$ and $3$. When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, so with this manner we know that our degrees are not to high. The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables. We can see that feature transformations significantly improve the test error compared to other models. However, dummy encoding doesn't seem to more efficient, and so our final model will be the seventh.

@@

We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and was less efficient than ridge regression. Separation of the data into clusters was essential in order to decrease the test error. Unfortunately, dummy encoding didn't help. However using polynomial basis helped to decrease the bias. Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher. We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient. The best model we have is the seventh one, and we estimate the \textit{RMSE} of the prediction to be around $97$ with a standard deviation of $35$. Finally, we think that our model is improvable, however we couldn't try other things in order to improve the estimation error, due to a lack of time.

@@

Since our matrix is full-rank using logistic regression was appropriate. Because logistic regression is more efficient in term of complexity and that we don't need to penalize some betas, we report only the results using this method. However, using penalized logistic regression gave us similar results.

@@

The fourth model consists in using dummy encoding for the categorical variable. This times, the error was improved and the variance has diminished. The next model (fifth) adds a polynomial basis of degree 2. We can observe the the variance as increased, we have some outliers (five in total, there might be hidden due to the legend) and the error has increased. We think that the bias is better but due to its high variance, we consider it as a bad model.

@@

The two last models (eighth and ninth) use another transformation where negative values weren't transformed into positive one, holding them with the formula $-\sqrt[4]{-x}$ and try using also a polynomial basis. However, the error wasn't as good as the seventh model, surely because the transformed feature look like less Gaussian than the transformation using the absolute values. We can also see that the variance seems also higher.
Our final model will be the seventh.

@@

In hope to minimize the error, we tried different feature transformations. The first try was add polynomial of them self as $X = [X X^2]$. With the data not transformed into Gaussian like distribution, we didn't get an improvement (even with degree 3). We then try to transform the features which look like a Poisson distribution in order to transform them into a Gaussian-like distribution. We first use the formula $\sqrt[4]{|x|}$ which gave us good results even if we transformed the negative values into positive. Mixed with polynomial of features at degree 2 allowed us to have an important improvement for the error. We also try to keep the negative values, using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help. We think the reason is first that distribution was more a superposition of two Gaussian. Using a degree greater than 2 didn't improve the error, surely because our bias is not too high.

@@

We analyzed a classification data set with different methods and concluded that logistic regression was appropriate because the input matrix was full-rank. Moreover, using the penalized logistic regression gave us similar results. Using normalization was essential as well as transforming the output in the range $[0,1]$. Adding categorical variables with dummy encoding helps. The key of our model was to transform features which look like Poisson distribution into a Gaussian like distribution using $\sqrt[4]{|x|}$ and using a polynomial basis of degree 2. We think that our model is improvable due to outliers because we couldn't find such outliers in all dimensions (or most of the dimensions). We estimate the mean and standard deviation of the prediction, using the \textit{RMSE},\textit{0-1 loss} and \textit{logarithmic loss} respectively to be around $0.329$, $0.109$ and $0.509$ for the mean, and $0.028$, $0.018$, $0.018$for the standard deviation .

>>>>

Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable. Moreover, the results obtained were less good compared to the ridge regression method, and so we will not report those. The method using the least squares with gradient descent is appropriated as well as ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix. It is also faster and a little bit more efficient than least squares using gradient descent and so, only results obtained with this latter will be reported.
\vspace{-0.5em}
\subsection{Evaluation methods}

@@

The sixth model consists of using dummy encoding for the categorical variables. Unfortunately, we didn't get major improvements. We think that the reason why dummy encoding didn't help is because we don't have a high bias and so, having more features doesn't help.

@@

For the seventh model, the degrees of the polynomial for the clusters are $3,6$ and $3$. When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, so with this manner we know that our degrees are not to high. The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables. We can see that feature transformations significantly improves the test error compared to other models. The reason is certainly because we had a high bias. However, dummy encoding doesn't seem to more efficient. Our final model will be the seventh with lambdas $\lambda_1 = 0.0369, \lambda_2 = 1.8303e^{-4}, \lambda_3 = 0.0053$.

@@

We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was too slow and less efficient than ridge regression. Separation of the data into clusters was essential in order to decrease the test error. Unfortunately, dummy encoding didn't help. However using polynomial basis helped to decrease the bias. Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher. We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps until a certain point. The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.96$ with a standard deviation of $45.59$. However, we should take into consideration outliers, which gives us a \textit{RMSE} of $407.55$ with a standard deviation of $229.34$. Finally, we think that our model is improvable, however we couldn't try other things to improve the estimation error, due to a lack of time.

@@

Since our matrix is full-rank, using logistic regression was appropriate. Because logistic regression is more efficient in term of complexity and that we don't need to penalize some betas, we report only the results using this method. However, using penalized logistic regression gave us similar results.

@@

The fourth model consists in using dummy encoding for the categorical variables. This time the error was improved and the variance has diminished as well as certainly the bias due to the larger set of features. The next model (fifth) adds a polynomial basis of degree 2. We can observe the the variance as increased, we have some outliers (five in total, hidden by the legend) and the error has increased. We think that the bias is better but due to its high variance, we consider it as a bad model.

@@

The two last models (eighth and ninth) use another transformation where negative values weren't transformed into positive one with the formula $-\sqrt[4]{-x}$ and try using also a polynomial basis. However, the error wasn't as good as the seventh model, surely because the transformed feature look like less Gaussian than the transformation using the absolute values. We can also see that the variance seems also higher.
Our final model will be the seventh with a learning step $\alpha=0.001$.

@@

In hope to minimize the error, we tried different feature transformations. We first added polynomial of them self as $X = [X X^2]$. With the data not transformed into Gaussian like distribution, we didn't get an improvement (even with degree 3). We then tried to transform the features which look like a Poisson distribution in order to transform them into a Gaussian-like distribution. We first use the formula $\sqrt[4]{|x|}$ which gave us good results even if we transformed the negative values into positive. Mixed with polynomial of features at degree 2 allowed us to have an important improvement for the error. We also try to keep the negative values using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help. We think the reason is first that distribution was more a superposition of two Gaussian. Using a degree greater than 2 didn't improve the error, surely because our bias is not too high.

@@

We analyzed a classification data set with different methods and concluded that logistic regression was appropriate because the input matrix was full-rank. Moreover, using the penalized logistic regression gave us similar results. Using normalization was essential as well as transforming the output in the range $[0,1]$. Adding categorical variables with dummy encoding helps. The key of our model was to transform features which look like Poisson distribution into a Gaussian like distribution using $\sqrt[4]{|x|}$ and using a polynomial basis of degree 2. We think that our model is improvable due to outliers because we couldn't find such in most of the dimensions. We estimate the mean and standard deviation of the prediction, using the \textit{RMSE},\textit{0-1 loss} and \textit{logarithmic loss} respectively to be around $0.329$, $0.109$ and $0.509$ for the mean, and $0.028$, $0.018$, $0.018$ for the standard deviation .
